Limb-darkening and exoplanets II: Choosing the Best Law for Optimal Retrieval of Transit Parameters

### 2016-01-15 first_response_to_referee.txt

 We thank the referee for his/her careful reading of the manuscript. 
 You will find below our detailed response to the referee's comments 
 as indented text after each of his/her comments.

 Besides the changes made in response to the referee's comments we have:

     - Corrected minor typos on the manuscript.

     - Changed the title of the manuscript to reflect the additional study 
       proposed by the referee on the precision of the retrieved transit 
       parameters, which implies we now not only help to decide the law 
       that provide minimum bias, but also the one that provides the 
       best precision. This in turn means that we help to decide the 
       optimal law, which is a statistical jargon for "lowest bias, 
       lowest variance" model.

     - Added extra text on the Abstract, Introduction and Section 3 to reflect 
       these changes in the study. This includes a change of the name of Section 3, 
       and an additional sub-section explaining the procedures used to study the 
       bias/variance trade-off.

     - Modified the Acknowledgments.

@@

########################################################################################
>> My main suggestion is that the paper would benefit from a more quantitative 
>> description of the trade-off between precision and accuracy. This work indicates 
>> that fitting for two limb darkening coefficients with a judiciously chosen model 
>> can significantly reduce bias. In practice, however, many light curves are not 
>> sufficiently precise to tightly constrain more than one limb darkening parameter. 
>> Adding an additional degree of freedom with a second limb darkening coefficient 
>> may cause the uncertainty on other important parameters, particularly the transit 
>> depth, to blow up. The resulting loss in accuracy may not be worth the improvement 
>> in precision due to choosing a two-parameter model over fitting for  a linear limb 
>> darkening coefficient.
>>
>> One way to address this question is to give guidelines about how precise a 
>> light curve must be to merit adding a second limb darkening parameter to the 
>> model. This could be discussed for a few select examples (e.g., a hot Jupiter 
>> transiting a Sun-like star).  I leave it up to the authors whether/how to address 
>> this point.

We thank the referee for this very insightful suggestion. We have made a whole suit of 
simulations which exactly answers this question, and provide code so future researchers 
can decide which is the best law both in terms of bias and precision for their specific 
application; in this way, researchers can decide the optimal limb-darkening law for the 
lightcurve at hand. These simulations are shown and discussed both on Section 3.4 and 
Section 4.

@@

>> It would also be helpful to investigate how much bias is introduced by using a linear 
>> limb darkening law, since this model is commonly used in practice. Again, the authors 
>> can address this at their discretion.

We thank the referee for making this suggestion. We have added simulations that provide 
the results for the linear law. As expected, the bias is larger than for any of the other 
laws, and this is now discussed both in the abstract and in Sections 3 and 4.

### 2016-01-15 which_law_2_use.tex
Finished response to the referee. Added two new simulations: one with the linear law, and one on the bias/variance trade-off. Edits and comments are welcome.

\title[Choosing the best law for minimum bias retrieval of transit parameters]
{Limb-darkening and exoplanets II: Choosing the Best Law for Minimum Bias Retrieval of Transit Parameters}
\author[Espinoza \& Jord\'an ]{
N\'estor Espinoza$^{1,2}$\thanks{E-mail:nespino@astro.puc.cl}, 
Andr\'es Jord\'an$^{1,2}$\thanks{E-mail:ajordan@astro.puc.cl}\\ 
$^{1}$Instituto de Astrof\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~na Mackenna 
4860, Santiago, Chile\\
$^{2}$Millennium Institute of Astrophysics, Santiago, Chile\\
}
\begin{document}

@@

\begin{abstract}
With the advent of very precise measurements of exoplanet transit lightcurves both from ground and space based observatories, it has been 
shown that it is now possible to fit the limb-darkening coefficients in the transit-fitting procedure rather than fix them to theoretical values. 
This strategy has been shown to give better results as fixing the coefficients to theoretical values can give rise to important yet unknown systematic 
errors, which directly impact the physical properties of the system derived from such lightcurves such as the exoplanet's radius, scaled 
semi-major axis and inclination. However, studies of the effect of limb darkening assumptions on the retrieved parameters 
have mostly focused on the widely used quadratic limb-darkening law, leaving out other proposed laws that are in fact better descriptions 
of model intensity profiles, which in principle could diminish the bias even further. In this work, we show that laws such as the logarithmic, 
square-root and three-parameter do a better job in general at retrieving the mentioned parameters from transit lightcurves than the widely 
used quadratic law and therefore we recommend using those instead. We also provide code to guide the decision of when to use each of 
the laws,  which we note has a dependence on both stellar and transit parameters. Finally, we demonstrate that the so-called exponential 
law is non-physical as it typically produces negative intensities close to the limb and should therefore not be used.
\end{abstract}

@@

In this work, we extend the work of EJ15 by testing how well these non-standard laws perform at retrieving transit parameters from transit lightcurves, and compare them to 
the retrievals done by the standard quadratic limb-darkening law. We follow \cite{kipping2013} in order to derive an efficient sampling scheme 
for the standard form of the logarithmic and exponential laws, and show that the latter is actually \textit{always} non-physical in the sense proposed by 
\cite{kipping2013}. Then, we simulate transit lightcurves and perform a simulation study similar to the one done in EJ15 using the derived sampling schemes 
for the logarithmic law and the sampling schemes detailed in \cite{kipping2013} for the other two-parameter laws and \cite{kipping2015} for the three-parameter law. We use our simulations to provide guidance as to when a particular law should be preferred.

This work is organised as follows. In \S2, we revisit the logarithmic and exponential 
laws in order to derive an efficient informative sampling strategy following the methods in \cite{kipping2013} for the standard forms of these laws. 
In \S3 we use those results in addition to the ones published by \cite{kipping2013} and \cite{kipping2015} in order to simulate transit lightcurves to test 
how well these laws perform in retrieving transit parameters. In \S4 we present a discussion and conclusions of our work.

@@

\section{Biases in transit parameters using two-parameter laws}

Having dealt with the sampling of limb-darkening coefficients from the 
logarithmic and exponential laws, we now turn to the study of the 
biases introduced on retrieved transit parameters in the transit fitting process 
introduced by using different limb-darkening laws. We perform a similar study as the one done in EJ15, but now 
(1) using the \texttt{batman} transit code described in \cite{batman2015}, (2) using three different two-parameter 
laws: the quadratic, logarithmic and square-root laws, and the three-parameter law and (3) smaller number of steps in 
the planet-to-star radius ratio, $p=R_p/R_*$, in order to illustrate the evolution of the biases with this parameter 
in a cleaner way: small ratios ($p=0.01$), medium ratios ($p=0.07$) and large ones ($p=0.13$).

As in EJ15, we simulate transit lightcurves for planets orbiting host stars with solar metallicity, 
gravity and microturbulent velocity, and $T_{\rm eff}$ between 3500 K and 9000 K, with the mentioned planet-to-star 
radius ratios, values of the scaled semi-major axis of $a_R = a/R_* = \{3.27, 3.92, 4.87, 6.45, 9.52, 18.18, 200\}$ and different 
impact parameters. The transits are simulated using a non-linear law in order to emulate ``real" intensity profiles, 
with the coefficients generated using the ATLAS models and the same methods as the ones used in EJ15 with the 
{\em Kepler} bandpass. $1000$ in-transit points and $400$ out-of-transit points were simulated per lightcurve, with the initial 
times perturbed by a small amount in order for the results to be independent of the exact points at which the lightcurve is sampled 
(which anyways our simulations show has a negligible effect given the large amount of sampled points). The transits were then fitted 
using free limb-darkening coefficients.

For the simulations, the applied strategies to sample physically plausible coefficients from two-parameter laws 
are as follows, all of which use numbers $q_1 \in (0,1)$ and $q_2 \in (0,1)$:
\begin{itemize}
\item For the quadratic limb-darkening law, we fit for the numbers $q_1 = (u_1 + u_2)^2$ and $q_2 = u_1/2(u_1+u_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $u_1 = 2\sqrt{q_1}q_2$ and $u_2 =\sqrt{q_1}(1-2q_2)$, 
which are derived in \cite{kipping2013}.
\item For the square-root law, we fit for the numbers $q_1 = (s_1 + s_2)^2$ and $q_2 = s_2/2(s_1+s_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $s_1 = \sqrt{q_1}(1-2q_2)$ and $s_2 =2\sqrt{q_1}q_2$, 
which are derived in \cite{kipping2013}.
\item For the logarithmic law, we fit for the numbers $q_1 = (1-l_2)^2$ and $q_2 = (1-l_1)/(1-l_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $l_1 = 1-\sqrt{q_1}q_2$ and $l_2=1-\sqrt{q_1}$, 
which are derived in \S 2.1.
\end{itemize}
For the three-parameter law, we use the formalism and code described in \cite{kipping2015}.  The code used to generate our simulations is available
in GitHub\footnote{\url{http://www.github.com/nespinoza/ld-exosim/}}, and can be used to perform analyses in a finer grid than the ones 
published here.

\subsection{The case of central transits}
\begin{figure*}
\includegraphics[scale=0.8]{bias_transit_params/all_laws/b_0.0/simulation_b_00_float.eps}
\caption{Results of fitting transit lightcurves as described in \S 3.1 with the limb-darkening coefficients as free parameters. The size of the points is proportional to the the input values of $p$.}
\label{central_sims}
\end{figure*}

The results of the simulations for central transits (i.e., $b=0$) are shown in Figure~\ref{central_sims}, with the sizes of the points proportional to the 
input values of $p$. The upper panels show the biases 
introduced on the planet-to-star radius ratio, $p$, and the lower panels show the biases introduced on the scaled semi-major axes, $a_R$. From left to right, 
the panels show the biases introduced by using a quadratic law, a logarithmic law, a square-root law and a three-parameter law. The dependence with the input values of $a_R$ are 
not shown as the bias seems to be almost independent of this parameter in this case (this can be also seen by looking at the lower panels of Figure~10 in EJ15, which 
give the same results as the ones shown here for the quadratic law. We note that this implies good agreement between the simulations done with our implementation of 
the \cite{ma2002} codes for the transit light curve generation and the 
\texttt{batman} code).

As can be seen from the figure, different limb-darkening laws have different behaviours in retrieving $p$ and $a_R$, with a dependence on the input value of 
$p$, and also depending on the temperature of the host star. First, it is interesting to note that the logarithmic law does a very good job at retrieving 
the parameters of small planet-to-star radius ratios, even at the level of the three-parameter law for stars with $T_\textnormal{eff}<6000$ K, and also for medium 
and large planet-to-star radius ratios with host-stars with $T_\textnormal{eff}<4000$ K (which makes this law a good choice for present and future transit 
detection and characterisation around M-dwarfs). The quadratic law, on the other hand, is only the best among the options for host-stars between 
$4250<T_\textnormal{eff}<5500$ K, with the logarithmic law being the best in the $5500<T_\textnormal{eff}<6250$ K range, and the square-root law 
being the best for planets around host stars with temperatures between $6250<T_\textnormal{eff}<8000$ K. Hotter than this, the best option in terms 
of accuracy is the three-parameter law.

It is interesting to emphasise that although the biases introduced by the quadratic law might seem to be small, as mentioned in the introduction and in 
EJ15 they are significant for several candidate and confirmed exoplanets. Choosing the right law  can help to achieve the same 
level of accuracy than the level of precision that {\em Kepler} achieves for most planets. For example, accuracies on the order of $\sim 0.01\%$ can 
be achieved for $p$ with a careful selection of a two-parameter law, and exoplanets with precisions better than that according to a query to the Nasa Exoplanet 
Archive\footnote{Query done on 29/09/2015.} form only $0.58\%$ of all candidate exoplanets and $1.2\%$ of the confirmed exoplanets. For the remaining 
exoplanets requiring a better accuracy, a new law might have to be created, as this precision is also the limit of the three-parameter law (which has a precision 
on $p$ around $\sim 0.01\%$ for stars with temperatures larger than $T_\textnormal{eff}=6000$ K). Studying new limb-darkening laws, however, 
is out of the scope of the present work.

\subsection{The case of non-central transits}
\begin{figure*}
\includegraphics[scale=0.8]{bias_transit_params/all_laws/b_0.3/simulation_b_03_float.eps}
\caption{Results of fitting transit lightcurves as described in \S 3.1 with the limb-darkening coefficients as free parameters, but for low impact 
parameter transits ($b=0.3$). The size of the points represent the input values of $p$, while their color represent the input values of $a_R$.}
\label{small_b_sims}
\end{figure*}

\begin{figure*}
\includegraphics[scale=0.8]{bias_transit_params/all_laws/b_0.8/simulation_b_08_float.eps}
\caption{Results of fitting transit lightcurves as described in \S 3.1 with the limb-darkening coefficients as free parameters, but for high impact 
parameter transits ($b=0.8$). The size of the points represent the input values of $p$, while their color represent the input values of $a_R$.}
\label{big_b_sims}
\end{figure*}

Figures~\ref{small_b_sims} and \ref{big_b_sims} show the results for low ($b=0.3$) and high ($b=0.8$) input impact parameters respectively. In these cases, 
we see that the biases are in general worse for smaller values of $p$ and $a_R$, i.e., for a given stellar radius, for smaller, close-in planets. We can 
also see that, again, a careful selection of the law to be used can allow one to minimize the biases on the retrieved transit parameters. For example, if 
the objective were to obtain $p$ with minimum bias for a low impact-parameter transit, then for host stars colder than $4000$ K the best option is to use 
either the logarithmic or the three-parameter law, which show biases on the order of $\sim 0.1\%$. In contrast, using the quadratic law for transits around 
those host stars the biases introduced on this parameter are on the order of $\sim 0.75\%$, i.e., almost an order of magnitude larger.

@@

@@

In this work we have explored the different biases on retrieved transit parameters incurred by the usage of different limb-darkening laws. In particular, we 
explored the biases introduced by using three two-parameter laws (the quadratic, logarithmic and square-root laws) and the three-parameter law of \citet{sing2009}. 
We showed that a careful selection of the limb-darkening law to be used in a given dataset can allow one to retrieve minimum bias estimators for the transit 
parameters from transit lightcurves. In general, as shown in \S3, two-parameter laws have ranges in which they perform at the same level or even better 
than the three-parameter law and, because of this, we encourage their use. We also discourage the usage of the exponential law which, as shown in \S2.2 is 
unphysical, and can lead to numerical errors if used on real data where sampling of points close to the limb (where the law diverges) is not uncommon.

The question of when to use each of those laws is an important one. In particular, we have explored the relation between the real, underlying 
transit parameters, the stellar host properties and the bias on the retrieved transit parameters. We have stated no specific rule regarding when to use each 
law, but rather provided only guidelines based on our simulations. This is a choice made on purpose, as the selection of which law to use has to be done in a case-by-case basis, reason for which we provide a code with which one can obtain the results of the same simulations for specific 
cases and with different resolutions\footnote{\url{http://www.github.com/nespinoza/ld-exosim/}}. It is important to note that in this work we have used the 
{\em Kepler} bandpass to study the relation between the biases on the retrieved transit parameters and the stellar host properties, but different bandpasses should give 
rise to  different functional relations because they give rise to different model intensity profiles, which are used in those simulations to approximate the 
expected intensity profile and the adequacy of each limb-darkening law at following it. Limb-darkening coefficients have to be calculated for each 
particular case, which is straightforward to do using the algorithms described in EJ15\footnote{\url{http://www.github.com/nespinoza/limb-darkening/}}.

>>>>

\title[Choosing the best law for optimal retrieval of transit parameters]
{Limb-darkening and exoplanets II: Choosing the Best Law for Optimal Retrieval of Transit Parameters}
\author[Espinoza \& Jord\'an ]{
N\'estor Espinoza$^{1,2}$\thanks{E-mail:nespino@astro.puc.cl}, 
Andr\'es Jord\'an$^{1,2}$\thanks{E-mail:ajordan@astro.puc.cl}\\ 
$^{1}$Instituto de Astrof\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~na Mackenna 
4860, Santiago, Chile\\
$^{2}$Millennium Institute of Astrophysics, Santiago, Chile\\
}
\begin{document}

@@

\begin{abstract}
With the advent of very precise measurements of exoplanet transit lightcurves both from ground and space based observatories, it has been 
shown that it is now possible to fit the limb-darkening coefficients in the transit-fitting procedure rather than fix them to theoretical values. 
This strategy has been shown to give better results as fixing the coefficients to theoretical values can give rise to important yet unknown systematic 
errors, which directly impact the physical properties of the system derived from such lightcurves such as the exoplanet's radius, scaled 
semi-major axis and inclination. However, studies of the effect of limb darkening assumptions on the retrieved parameters 
have mostly focused on the widely used quadratic limb-darkening law, leaving out other proposed laws that are either simpler (e.g., the linear law) or 
in fact better descriptions of model intensity profiles, which in principle could diminish the bias even further. In this work, we show that laws such as the 
logarithmic, square-root and three-parameter do a better job in general at retrieving the mentioned parameters from transit lightcurves both in terms of 
bias and precision than the widely used quadratic and linear laws and therefore we recommend using those instead. We also provide code to guide the 
decision of when to use each of these laws,  which we note has a dependence on both stellar and transit parameters. Finally, we demonstrate that the 
so-called exponential law is non-physical as it typically produces negative intensities close to the limb and should therefore not be used.
\end{abstract}

@@

In this work, we test how well these non-standard laws perform at retrieving transit parameters from transit lightcurves. We follow \cite{kipping2013} in order to 
derive an efficient sampling scheme for the standard form of the logarithmic and exponential laws, and show that the latter is actually \textit{always} non-physical 
in the sense proposed by \cite{kipping2013}. Then, we simulate transit lightcurves and perform simulations studies which extend the one done in EJ15 in several ways. 
First, we study not only the accuracy problem but also the precision issue on the retrieved parameters. Second, we study the mentioned non-standard limb-darkening laws by 
using the derived sampling schemes for the logarithmic law derived in this work, the sampling schemes detailed in \cite{kipping2013} for the other two-parameter laws 
and \cite{kipping2015} for the three-parameter law. In addition, we also study the performance of the linear law, which has been used in recent years by several authors 
to parametrize the limb-darkening effect. We use our simulations to provide guidance as to when a particular law should be preferred, and provide code which, given the 
precision of a given transit lightcurve and a set of transit and stellar parameters, helps to decide which law is the optimal (i.e., with the best bias/variance trade-off) one to use.

This work is organised as follows. In \S2, we revisit the logarithmic and exponential laws in order to derive an efficient informative sampling strategy following the methods 
in \cite{kipping2013} for the standard forms of these laws. In \S3 we use those results in addition to the ones published by \cite{kipping2013} and \cite{kipping2015} in order 
to simulate transit lightcurves to test how well these laws perform in retrieving transit parameters both in terms of accuracy and precision. In \S4 we present a discussion and 
conclusions of our work.

@@

\section{Optimal selection of limb-darkening laws}
Having dealt with the sampling of limb-darkening coefficients from the 
logarithmic and exponential laws, we are now ready to turn to the study of the performance 
of the most popular limb-darkening laws (i.e., the linear, quadratic, logarithmic, square-root and 
three-parameter laws) both in terms of bias and precision in the 
retrieval of transit parameters. In particular, we focus on the performance of the retrieval of the planet-to-star 
radius ratio, $p=R_p/R_*$, the scaled semi-major axis, $a_R = a/R_*$ and the inclination, $i$, from transit lightcurves. 
In this section, we first study the biases introduced by using different limb-darkening laws, and then we study the 
bias-variance trade-off between different laws.

\subsection{Biases in the retrieval of transit parameters}
In order to study the biases introduced by different parametrizations of the limb-darkening effect, we perform a similar study 
as the one done in EJ15, but now (1) using the \texttt{batman} transit code described in \cite{batman2015}, (2) using the 
linear law, three different two-parameter laws (the quadratic, logarithmic and square-root laws) and the three-parameter 
law and (3) using a smaller number of steps in the planet-to-star radius ratio, in order to illustrate the evolution of the biases 
with this parameter in a cleaner way: we choose to study small ratios ($p=0.01$), medium ratios ($p=0.07$) and large ones 
($p=0.13$).

As in EJ15, we simulate transit lightcurves for planets orbiting host stars with solar metallicity, 
gravity and microturbulent velocity, and $T_{\rm eff}$ between 3500 K and 9000 K, with the mentioned planet-to-star 
radius ratios, values of the scaled semi-major axis of $a_R = \{3.27, 3.92, 4.87, 6.45, 9.52, 18.18, 200\}$ and different 
impact parameters. The transits are simulated using a non-linear law in order to emulate ``real" intensity profiles, 
with the coefficients generated using the ATLAS models and the same methods as the ones used in EJ15 with the 
{\em Kepler} bandpass. As in this previous study, $1000$ in-transit points and $400$ out-of-transit points were simulated per 
lightcurve, with the initial times perturbed by a small amount in order for the results to be independent of the exact points at 
which the lightcurve is sampled (which anyways our simulations show has a negligible effect given the large amount of sampled 
points). The transits were then fitted using free limb-darkening coefficients with the different laws mentioned above.

For the simulations, the applied strategies to sample physically plausible coefficients from two-parameter laws 
are as follows, all of which use numbers $q_1 \in (0,1)$ and $q_2 \in (0,1)$:
\begin{itemize}
\item For the quadratic limb-darkening law, we fit for the numbers $q_1 = (u_1 + u_2)^2$ and $q_2 = u_1/2(u_1+u_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $u_1 = 2\sqrt{q_1}q_2$ and $u_2 =\sqrt{q_1}(1-2q_2)$, 
which are derived in \cite{kipping2013}.
\item For the square-root law, we fit for the numbers $q_1 = (s_1 + s_2)^2$ and $q_2 = s_2/2(s_1+s_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $s_1 = \sqrt{q_1}(1-2q_2)$ and $s_2 =2\sqrt{q_1}q_2$, 
which are derived in \cite{kipping2013}.
\item For the logarithmic law, we fit for the numbers $q_1 = (1-l_2)^2$ and $q_2 = (1-l_1)/(1-l_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $l_1 = 1-\sqrt{q_1}q_2$ and $l_2=1-\sqrt{q_1}$, 
which are derived in \S 2.1.
\end{itemize}
For the three-parameter law, we use the formalism and code described in \cite{kipping2015}, while for the linear law we simply 
sample coefficients between $0$ and $1$, which give intensity profiles which are both always positive and which show a strictly 
decreasing intensity profiles from center to limb.  The code used to generate our simulations is available in GitHub\footnote{\url{http://www.github.com/nespinoza/ld-exosim/}}, 
and can be used to perform analyses in a finer grid than the ones published here.

\subsubsection{The case of central transits}
\begin{figure*}
\includegraphics[scale=0.67]{bias_transit_params/all_laws/b_0.0/simulation_b_00_float.eps}
\caption{Results of fitting transit lightcurves as described in \S 3.1 with the limb-darkening coefficients as free parameters. Note that the scale shown for the linear law is different 
than the one used in the figures for the other laws. The size of the points is proportional to the the input values of $p$.}
\label{central_sims}
\end{figure*}

The results of the simulations for central transits (i.e., $b=0$) are shown in Figure~\ref{central_sims}, with the sizes of the points proportional to the 
input values of $p$. The upper panels show the biases introduced on the planet-to-star radius ratio, $p$, and the lower panels show the biases introduced 
on the scaled semi-major axes, $a_R$. From left to right, the panels show the biases introduced by using a linear law, a quadratic law, a logarithmic law, a 
square-root law and a three-parameter law. The dependence with the input values of $a_R$ are not shown as the bias seems to be almost independent of 
this parameter in this case (this can be also seen by looking at the lower panels of Figure~10 in EJ15, which give the same results as the ones shown here for 
the quadratic law. We note that this implies good agreement between the simulations done with our implementation of the \cite{ma2002} codes for the transit light 
curve generation and the \texttt{batman} code).

As can be seen from the figure, different limb-darkening laws have different behaviours in retrieving $p$ and $a_R$, with a dependence on the input value of 
$p$, and also depending on the temperature of the host star. First, it is interesting to note that the linear law is by far the worst of the laws in terms of bias. As 
expected from its very simple form, the biases on the retrieved transit parameters are larger than the ones observed on the two and three-parameter laws for 
almost all the stellar temperatures except for $T_\textnormal{eff}\sim 4500$, where the profile seems to be a decent (yet poorer than the ones provided by the 
other laws) approximation to the real intensity profile. A second feature to note on our results is that the logarithmic law does a very good job at retrieving the 
parameters of small planet-to-star radius ratios, even at the level of the three-parameter law for stars with $T_\textnormal{eff}<6000$ K, and also for medium 
and large planet-to-star radius ratios with host-stars with $T_\textnormal{eff}<4000$ K (which makes this law a good choice for present and future transit 
detection and characterisation around M-dwarfs). The quadratic law, on the other hand, is only the best among the options for host-stars between 
$4250<T_\textnormal{eff}<5500$ K, with the logarithmic law being the best in the $5500<T_\textnormal{eff}<6250$ K range, and the square-root law 
being the best for planets around host stars with temperatures between $6250<T_\textnormal{eff}<8000$ K. Hotter than this, the best option in terms 
of accuracy is the three-parameter law.

It is interesting to emphasise that although the biases introduced by the quadratic law might seem to be small, as mentioned in the introduction and in 
EJ15 they are significant for several candidate and confirmed exoplanets. Choosing the right law can help to achieve the same 
level of accuracy than the level of precision that {\em Kepler} achieves for most planets. For example, accuracies on the order of $\sim 0.01\%$ can 
be achieved for $p$ with a careful selection of a two-parameter law, and exoplanets with precisions better than that according to a query to the Nasa Exoplanet 
Archive\footnote{Query done on 29/09/2015.} form only $0.58\%$ of all candidate exoplanets and $1.2\%$ of the confirmed exoplanets. For the remaining 
exoplanets requiring a better accuracy, a new law might have to be created, as this is also the limit of the three-parameter law (which has an accuracy 
on $p$ around $\sim 0.01\%$ for stars with temperatures larger than $T_\textnormal{eff}=6000$ K). Studying new limb-darkening laws, however, 
is out of the scope of the present work.

\subsubsection{The case of non-central transits}
\begin{figure*}
\includegraphics[scale=0.67]{bias_transit_params/all_laws/b_0.3/simulation_b_03_float.eps}
\caption{Results of fitting transit lightcurves as described in \S 3.1 with the limb-darkening coefficients as free parameters, but for low impact 
parameter transits ($b=0.3$). Note that the scale shown for the linear law is different than the one used in the figures for the other laws. The 
size of the points represent the input values of $p$, while their color represent the input values of $a_R$.}
\label{small_b_sims}
\end{figure*}

\begin{figure*}
\includegraphics[scale=0.67]{bias_transit_params/all_laws/b_0.8/simulation_b_08_float.eps}
\caption{Results of fitting transit lightcurves as described in \S 3.1 with the limb-darkening coefficients as free parameters, but for high impact 
parameter transits ($b=0.8$). Note that the scale shown for the linear law is different than the one used in the figures for the other laws. The size 
of the points represent the input values of $p$, while their color represent the input values of $a_R$.}
\label{big_b_sims}
\end{figure*}

Figures~\ref{small_b_sims} and \ref{big_b_sims} show the results for low ($b=0.3$) and high ($b=0.8$) input impact parameters respectively. In these cases, 
we see that the biases are in general worse for smaller values of $p$ and $a_R$, i.e., for a given stellar radius, for smaller, close-in planets. We can 
also see that, again, the worst law to choose in terms of accuracy is the linear, which can give rise to biases on the order of $\sim 3\%$ for small planet-to-star 
radius ratios. However, it is evident from the results that a careful selection of the law to be used can allow one to minimize the biases on the retrieved transit 
parameters. For example, if the objective were to obtain $p$ with minimum bias for a low impact-parameter transit, then for host stars colder than $4000$ K 
the best option is to use either the logarithmic or the three-parameter law, which show biases on the order of $\sim 0.1\%$. In contrast, using the quadratic law 
for transits around those host stars can give rise to biases on the order of $\sim 0.75\%$, i.e., almost an order of magnitude larger.

@@

\subsection{The bias-variance trade-off}
From the results on the past sub-section, one can see that for a given set of stellar and transit parameters, a law can be chosen in order to minimise the 
biases in the retrieved transit parameters from the transit light curve. However, an interesting question is whether the precision of a given lightcurve can 
actually provide one with enough precision in order to use a more complex parametrization of the limb-darkening effect and, depending on the answer, what is 
the cost. Fitting a complex law to a noisy light curve will definitely lower the precision on the retrieved transit parameters due to the high flexibility of the 
model (i.e., the large parameter space that it can explore), while fitting a law that is too simple for a very precise lightcurve will most definitely lead to larger 
biases, due to the strictness of the chosen model. This problem is known as the bias-variance trade-off, i.e., the problem of choosing the right amount of bias 
and variance for a given dataset. Probabilistically, for a given estimator $\hat{\theta}$ of the parameter $\theta$, we will be interested in its variance (whose 
square-root defines the precision), $\textnormal{Var}[\hat{\theta}]$, while the bias of the estimator is defined as $\mathbb{E}[\hat{\theta}]-\theta$, where 
$\mathbb{E}[\cdot]$ defines the expectation value. In this subsection, we will 
be interested in the ratio of the \textit{absolute value} of the bias over the precision, which we define as
\begin{eqnarray*}
\textnormal{Bias/Precision} = \frac{|\mathbb{E}[\hat{\theta}]-\theta|}{\sqrt{\textnormal{Var}[\hat{\theta}]}}
\end{eqnarray*}

\begin{figure*}
\includegraphics[scale=0.9]{which_law_to_use/bias_variance_tradeoff.eps}
\caption{Results of our lightcurve fits to noisy lightcurves of a typical Hot-Jupiter in order to study the bias-variance trade-off of different limb-darkening laws on the planet-to-star radius 
ratio for lightcurves with $100$ in-transit points. The ratio of the bias to the square-root of the variance (i.e., the precision) of each estimate for different laws is 
shown as a function of the lightcurve noise level. The line of equal bias and precision is shown as a dashed gray line.}
\label{biasvariance}
\end{figure*}

Given that the study of this ratio is complex from a mathematical point of view in this case, we study it through simulations similar to the ones done in the past sub-section, 
but this time also adding different amounts of noise to the lightcurves. As a case study, we take that of a typical Hot-Jupiter ($p=0.1$, $a_R = 10$) around a Solar-type 
star ($T_\textnormal{eff}= 5500$ K) on a low impact-parameter orbit ($b=0.3$), and we focus specifically on the retrieval of the planet-to-star radius ratio, which is 
one of the most interesting parameters retrieved from transit lightcurves. We simulate a suite of lightcurves with noise levels ranging from $10$ ppm to 
$3000$ ppm with different amount of in-transit points; here, however, we show our results for $N=100$ in-transit points for clarity, which is a worst-case scenario for 
typical \textit{Kepler} lightcurves, and is a typical case for transit lightcurves obtained from ground-based observatories. For each case, $300$ lightcurves are generated, 
and all of them are fitted in the same way as it was done for their noiseless counterparts in the past sub-section. From the fit to each of the $300$ lightcurves on each case, 
the precision of the planet-to-star radius ratio is estimated by calculating the square-root of the variance of all the retrieved planet-to-star radius ratio of each of the $300$ lightcurves. 
The results of our simulations are shown on Figure \ref{biasvariance}, where we plot the ratio of the absolute value of the bias (estimated in the previous sub-section) over the precision 
(the square-root of the variance of the estimator) as a function of the noise in the simulated lightcurves.

In our results, the line of equal bias and precision marks the limit at which one should switch to a more complex law. For example, in our simulation, the linear law touches this line at 
$\sim 500$ ppm, which means that for more precise lightcurves the bias on the planet-to-star radius ratio is larger than the precision (e.g., for a lightcurve with a precision of $\sim 200$ 
ppm, the bias is $\sim 2$ times larger than the precision on the parameter) and, therefore, one should switch to a more complex law. To which law should one switch? From Figure \ref{biasvariance}, 
it can be seen that between $\sim 50-500$ ppm, the best law in terms of bias/variance trade-off is the square-root law which, among all the laws that are below the equal bias and precision line, 
provides the best precision. For lightcurve precisions better than $\sim 50$ ppm, it can be seen that the square-root law crosses this line and therefore another law must be used, which can be 
selected in the same way as before (in this case, between $\sim 30-50$ ppm, the optimal decision would be to use the logarithmic law). For this specific case, it can be seen that the 
three-parameter law is only optimal for precision better than $\sim 20$ ppm. Of course, if one increases the number of in-transit points, the precisions at which each law crosses the equal bias 
and precision line are lower. For example, we find that for $N=500$, the linear law touches this line at $\sim 800$ ppm, while for $N=1000$, 
the linear law touches this line at $\sim 1000$ ppm.

From our case study, it can be seen that the selection of the optimal limb-darkening law (i.e., the one that provides the best precision among the ones that avoid biases larger than it) can be 
estimated by simple simulation. We provide code at Github with which this can be easily done for a given set of transit and stellar parameters, along with a given lightcurve noise level and number of 
in-transit points.

@@

In this work we have explored the different biases and precisions on retrieved transit parameters incurred by the usage of different limb-darkening laws. In particular, we 
explored the biases introduced by using the linear law, three two-parameter laws (the quadratic, logarithmic and square-root laws) and the three-parameter law of \citet{sing2009}, 
and defined a way to select the optimal law for a given lightcurve precision and number of in-transit points. We showed that a careful selection of the limb-darkening law to be 
used in a given dataset can allow one to retrieve optimal estimators for the transit parameters from transit lightcurves. In general, as shown in \S3, two-parameter laws have ranges 
in which they perform at the same level or even better than the three-parameter law and, because of this, we encourage their use. We also discourage the usage of the exponential 
law which, as shown in \S2.2 is unphysical, and can lead to numerical errors if used on real data where sampling of points close to the limb (where the law diverges) is not uncommon.

The question of when to use each of those laws is an important one. In particular, we have explored the relation between the real, underlying 
transit parameters, the stellar host properties and the bias/variance on the retrieved transit parameters. We have stated no specific rule regarding when to use each 
law, but rather provided only guidelines based on our simulations. This is a choice made on purpose, as the selection of which law to use has to be done in a 
case-by-case basis, reason for which we provide code with which one can obtain the results of the same simulations performed on our work for specific 
cases and with different resolutions\footnote{\url{http://www.github.com/nespinoza/ld-exosim/}}. It is important to note that in this work we have used the 
{\em Kepler} bandpass to study the relation between the biases/variances on the retrieved transit parameters and the stellar host properties, but different bandpasses should give 
rise to  different functional relations because they give rise to different model intensity profiles, which are used in those simulations to approximate the 
expected intensity profile and the adequacy of each limb-darkening law at following it. Limb-darkening coefficients have to be calculated for each 
particular case, which is straightforward to do using the algorithms described in EJ15\footnote{\url{http://www.github.com/nespinoza/limb-darkening/}}.
